---
title: "Some Replications of HanesCarpenter(1999)"
author: "Hyunwoo Gu"
date: "5/7/2019"
output: pdf_document
---

# Disclamier

Here I replicated some results of _Hanes \& Carpenter(1999)_ for the subject named **DH**. 

\pagebreak
# Data \& Figures

## Latency distribution of DH 

The following data were read from the **Fig. 2**.

```{r}
Time = c(170, 180, 190, 200, 210, 220, 230, 240, 250, 260, 270, 280, 290, 300,  310,  320, 330, 340)
HC   = c(0.3, 2,   7,   16,  31,  48,  62,  70,  80,  86,  92,  94,  96,  97.5, 99.1, 99.6, NA, NA)
LC   = c(NA , 0.3, 1.2, 3,   8,   18,  27,  38,  52,  61,  72,  79,  84,  90,   96,   98,  99.1, 99.7)

x_ticks = c(150, 200, 300, 500, 1000)
y_ticks = c(0.1,0.5,1,2,5,10,20,30,40,50,60,70,80,90,95,98,99,99.5,99.9)
```

The following is the latency distribution plot in original scale. 

```{r}
plot(Time, HC, pch=16, main="Latency Distribution of DH", 
     xlab="Sccadic Latency(ms)", ylab="Cumulative probability")
points(Time, LC, pch=1)
legend("bottomright", legend=c("High Contrast", "Low Contrast"),
       pch=c(16,1), box.lty=0)
```

\pagebreak
## Reciprobit plot of Latency distribution of DH

This is the replication of **Fig. 2**.

```{r, warning = FALSE}
invTime   = 1/Time ## transfrom x axis into reciprocal scale
HC_probit = qnorm(HC/100); LC_probit = qnorm(LC/100) ## transfrom y axis into probit scale

## High contrast plot (without x/y labels)
plot(invTime, HC_probit, 
     xlim=rev(range(invTime)), pch=16, labels = FALSE,  ## plz ignore warnings... 
     main="Latency Distribution of DH", xlab="Sccadic Latency(ms)", ylab="Cumulative probability")

## Low contrast plot
points(invTime, LC_probit, xlim=rev(range(1/Time)), pch=1)
legend("bottomright", legend=c("High Contrast", "Low Contrast"),
       pch=c(16,1), box.lty=0)

## Add median line / labels
abline(h=qnorm(50/100)) # add median line
axis(1, at = 1/x_ticks, labels=x_ticks, tick=T) # add x labels
axis(2, at = qnorm(y_ticks/100), labels = y_ticks, tick=T) # add y labels
```

\pagebreak
## Inhibition function of DH

The following data were read from the **Fig. 3**.

```{r}
Delta      = c(0,   50,  90,  130, 170, 210)
HC_inhibit = c(.01, .08, .45, .82, .98, NA)
LC_inhibit = c(NA,  .02, .18, .54, .88, .99)
```

The following is plot of Inhibition function of DH

```{r}
plot(Delta, HC_inhibit, pch=16, 
     main="Inhibition function of DH", 
     xlab=paste(expression(delta), "(ms)"), ylab=expression(P(delta)))
lines(Delta, HC_inhibit)
points(Delta, LC_inhibit, pch=1)
lines(Delta, LC_inhibit)
```

\pagebreak
# Recovery of the original data

There is no raw data provided. Rather, only thing we can find is **cumulative distributions**. Therefore, here I recovered the original data through the method of **inverse transform sampling**, i.e. we can generate the original data $X$ from

$$
F_X^{-1} (U) 
$$

where $F_X(\cdot)$ denotes the CDF of $X$, and $U \sim Unif(0,1)$


## Recovery of control trial data of DH

First, let us start with approximating the `CDF` and the `inverse CDF` of the control trial data of DH. CDFs for `high constrast` and `low constrast` were approximated using **linear interpolation**. And the `inverse CDF` were approximated via **bisection method** and then **vectorized**. 

```{r}
cdf_HC = approxfun(Time, HC/100, method="linear", rule=2)
inv_cdf_HC = function(y){uniroot(function(x){cdf_HC(x)-y},interval=c(150,400))$root}
inv_cdf_HC = Vectorize(inv_cdf_HC)

cdf_LC = approxfun(Time, LC/100, method="linear", rule=2)
inv_cdf_LC = function(y){uniroot(function(x){cdf_LC(x)-y},interval=c(150,400))$root}
inv_cdf_LC = Vectorize(inv_cdf_LC)
```

Next, let us apporximate the original control trial data of DH from $F_X^{-1} (U)$. For convenience, assume the number of trials $N=1000$.

```{r}
U = runif(1000,0.01,0.99) 
HC_sample = inv_cdf_HC(U)
LC_sample = inv_cdf_LC(U)
```

Finally, let us do **sanity checks**. Note that the following **mean** values are close to what is given in **table 1**.

```{r}
mean(HC_sample); mean(LC_sample); 
```

Moreover, we can observer that the distribution of **RT** from the simulated data look **right-skewed**, which is characteristic of **response time** data. 

```{r}
par(mfrow=c(1,2))
plot(density(HC_sample), main="High constrast recovered")
plot(density(LC_sample), main="Low constrast recovered")
```


\pagebreak
# Estimation of SSRT

Estimation of **stop signal response time(SSRT)** can be done through two methods, **1) method of integration** and **2) method of difference of means**

## Method 1. integration

This method is trickier than the latter, since this requires **numerical integration**. 

First, let us approximate the above density functions in an explicit way, by linear interpolation.

```{r}
HC_dens_function = approxfun(density(HC_sample)$x, density(HC_sample)$y, method="linear", rule=2)
LC_dens_function = approxfun(density(LC_sample)$x, density(LC_sample)$y, method="linear", rule=2)
```

Second, let us construct the **loss function of SSRT estimation**, as follows. This evaluates the **sum of squares of errors** between the estimated **integrated value** and the **observed value**.  

```{r}
SSRT_estimate_loss_func = function(SSRT_hat, dens_fun, obj){
  integrated_val = vector(length=6)
  for (i in 1:6){
    lower_lim = 0
    upper_lim = Delta[i] + SSRT_hat 
    integrated_val[i] = integrate(dens_fun, lower_lim, upper_lim)$value
  }
  return(sum((obj-integrated_val)^2, na.rm=T))  ## SSE
}
```

Finally, let us find the **minimizer of the loss function**, the **SSRT estimate**. 

```{r}
optimize(SSRT_estimate_loss_func, interval=c(50,200), dens_fun=HC_dens_function, obj=HC_inhibit)
optimize(SSRT_estimate_loss_func, interval=c(50,200), dens_fun=LC_dens_function, obj=LC_inhibit)
```

We get the values around $126 \sim 128$. which are close to those in **table 2**.


## Method 2. difference of means

This method is more simpler(although its rationale is less intuitive), in that it only requires the **difference of means** between the **reaction time during control** and **inhibition function**.

We need to estimate the **mean of the inhibition function** using the method similar to above.

```{r}
cdf_HC_inhibit = approxfun(Delta, HC_inhibit, method="linear", rule=2)
cdf_LC_inhibit = approxfun(Delta, LC_inhibit, method="linear", rule=2)

inv_cdf_HC_inhibit = function(y){uniroot(function(x){cdf_HC_inhibit(x)-y},interval=c(0,250))$root}
inv_cdf_HC_inhibit = Vectorize(inv_cdf_HC_inhibit)

inv_cdf_LC_inhibit = function(y){uniroot(function(x){cdf_LC_inhibit(x)-y},interval=c(0,250))$root}
inv_cdf_LC_inhibit = Vectorize(inv_cdf_LC_inhibit)

U = runif(1000,0.05,0.95)
HC_inhibit_sample = inv_cdf_HC_inhibit(U)
LC_inhibit_sample = inv_cdf_LC_inhibit(U)
```

Next, let us find the difference between means.

```{r}
mean(HC_sample) - mean(HC_inhibit_sample);
mean(LC_sample) - mean(LC_inhibit_sample)
```

The values are not very distinct from those above (in method 1) and are also close to those in **table 2**.

\pagebreak
# Estimation of control trial parameters using LATER modeling

We can also do the **LATER** modeling, for the control trials. 

## Simulation of the control trial

We can simulate for the control trial using the following function.

```{r}
simul_control_trial = function(par, tau=60, theta=1, trials=2048, iter_max=200){
  mu = par[1]; sigma = par[2]
  activ_level = vector(length=trials)
  rt_result   = vector(length=trials)
  for (time in 1:iter_max){
    activ_level = activ_level + 0.01 * rnorm(trials, mean=mu, sd=sigma) ## accumulation of act. level
    rt_result[activ_level>theta] = tau + 10*time  ## unit : ms
    activ_level[activ_level>theta] = -Inf
  }
  return(rt_result)
}
```


## Estimation of $\mu_{go}$ and $\sigma_{go}$

We can construct the loss function as follows, which defines the `Kolmogorov-Smirnov` loss function. 

```{r}
Kolmo_Smir_loss_between_simul_and_observed = function(par, observed){
  simul_res = simul_control_trial(par) ## save the simulated data from given parameters
  ecdf_simul = ecdf(simul_res) ## empricial cdf from the simulated result
  sum(abs(ecdf_simul(Time)-observed/100), na.rm=T)
}
```

Next, we can find minimizer $(\mu, \sigma)$ of the above loss, using the finite-difference optimizer. 

```{r}
optim(par=c(6,1), Kolmo_Smir_loss_between_simul_and_observed, observed=HC) ## for high constrast
optim(par=c(6,1), Kolmo_Smir_loss_between_simul_and_observed, observed=LC) ## for low constrast
```

The estimation of **$\mu$** is quite accurate, in comparison with $table 3$. However, the estimation of **$\sigma$** is not quite accurate, which I cannot explain for the moment.


## Demonstration using the estimated values

For the demonstration, we can plot the **density** as follows. 

```{r, warning=FALSE}
par(mfrow=c(2,2))
simul_res_HC = simul_control_trial(c(6.09, 4.33))  
simul_res_LC = simul_control_trial(c(5.10, 2.99))  
plot(density(HC_sample, bw=5), main="HC recovered")
plot(density(simul_res_HC, bw=5), main="HC simulated") 
plot(density(LC_sample, bw=5), main="LC recovered")
plot(density(simul_res_LC, bw=5), main="LC simulated") 
```

Next, we can plot the **reciprobit function** as follows.

```{r, warning=FALSE}
ecdf_simul_HC = ecdf(simul_res_HC)
ecdf_simul_LC = ecdf(simul_res_LC)
simul_probit_HC = qnorm(ecdf_simul_HC(Time))
simul_probit_LC = qnorm(ecdf_simul_LC(Time))
par(mfrow=c(1,2))

## Recovered
plot(invTime, HC_probit, 
     xlim=rev(range(invTime)), pch=16, labels = FALSE,
     main="Recovered Latency", xlab="Sccadic Latency(ms)", ylab="Cumulative probability")
points(invTime, LC_probit, xlim=rev(range(1/Time)), pch=1)
legend("bottomright", legend=c("High Contrast", "Low Contrast"),
       pch=c(16,1), box.lty=0)
abline(h=qnorm(50/100)) # add median line
axis(1, at = 1/x_ticks, labels=x_ticks, tick=T) # add x labels
axis(2, at = qnorm(y_ticks/100), labels = y_ticks, tick=T) # add y labels

## Simulated
plot(invTime, simul_probit_HC, 
     xlim=rev(range(invTime)), pch=16, labels = FALSE,
     main="Simulated Distribution", xlab="Sccadic Latency(ms)", ylab="Cumulative probability")
points(invTime, simul_probit_LC, xlim=rev(range(1/Time)), pch=1)
legend("bottomright", legend=c("High Contrast", "Low Contrast"),
       pch=c(16,1), box.lty=0)
abline(h=qnorm(50/100)) # add median line
axis(1, at = 1/x_ticks, labels=x_ticks, tick=T) # add x labels
axis(2, at = qnorm(y_ticks/100), labels = y_ticks, tick=T) # add y labels
```
