---
title: "Homework #9"
author: "Hyunwoo Gu"
date: "4/16/2019"
output: 
  pdf_document:
    keep_tex: yes
---

# Using `R`

This approach saves all the samples in a trace into memory, and take `mean` and `variance` from the saved memory.

## Definition of trace generation and summary function

The following `traceMeanVar` function generates and summarises a trace.

```{r}
traceMeanVar = function(p, num_traces, N=100, legend=T){
  # step1. generate samples
  samples = NULL ## initialize history
  for (trace in 1:num_traces){
    X = cumsum(sample(c(1,-1), size=N, prob=c(p,1-p), replace=T)) ## sample trace
    samples = cbind(samples, X) ## accumulate traces to history
  }
  
  # step2. plot and print mean and variance
  colnames(samples) = sprintf("S%d",1:num_traces) ## label samples "S1","S2",...
  matplot(samples, col=1:num_traces, lty=1, type="l", xlab="N",ylab="S(N)") ## plot
  print(apply(samples, 2, function(x) {setNames(c(mean(x), var(x)), c("Mean", "Var"))}))
  if (legend==T) legend("bottomleft", legend=colnames(samples), col=1:num_traces, lty=1)
}
```

\pagebreak
## Problem 1.

For $p=0.5$ case, we can do as follows:

```{r}
set.seed(190416)
traceMeanVar(p=0.5, num_traces=3)
```


## Problem 2. 

For $p=0.55$ case, we can do as follows:

```{r}
set.seed(190416)
traceMeanVar(p=0.55, num_traces=3)
```


\pagebreak

# Using `C++`

Without the need to save a trace, this approach calculates `online` version of **mean** and **variance** as a new element in the trace is sampled. In this approach, `multithreading` is implemented, thus efficiently calculates each line of traces simultaneously. 

The following is a trace generating and summarising function.
